Automating G\"odel's Ontological Proof of God's Existence with Higher-order Automated Theorem Provers

### 2014-05-19 Papers/2014-ECAI/paper.tex
added minor modifications

In this paper we focus on the core aspect of our work related to AI: proof automation with HOL-ATPs (cf.~aspects (ii)--(iv)
above). % In addition, we present the detailed technical
% background, i.e., the embedding of \HOML in \HOL.  None of these
% aspects have been adressed in detail yet in any of our other existing 
% (short and partly non-reviewed) papers on the subject \cite{J30,W50,J29,J28}.
The particular contributions of this paper are as follows: 
% In Sec.~\ref{sec1} we present Scott's script \cite{ScottNotes} of
% G\"odel's proof, and we summarize the results of our
% recent automation of this script with HOL-ATPs. 
In Sec.~\ref{sec2} we present an elegant embedding of 
\HOML \cite{Gallin75,homl} in \HOL \cite{andrewsSEP,B5}. 
This background theory extends and
adapts the work as presented in \cite{J23,B9}.
In Sec.~\ref{sec3}, we present details on the
encoding of this embedding and of G{\"o}del's argument in the
concrete THF syntax \cite{J22} for \HOL, and we report on the
experiments we have conducted with HOL-ATPs.  The main findings of
these experiments are summarized in Sec.~\ref{sec4},
% : in addition to
% automating G{\"o}del's proof script we have automated also some further
% results (e.g.~modal collapse), and we have studied
% the argument under modified logical settings (varying vs.~constant domain semantics).
% (by switching from constant
% domain semantics to varying domain semantics)
Related and future work is addressed in Sec.~\ref{sec5}, and the paper
is concluded in Sec.~\ref{sec6}.
%  where we also discuss potential
% implications towards a \emph{computational theoretical
%   philosophy}.
None of the above aspects have been addressed (at least
not in depth) in any of our other existing (short and partly
non-reviewed) publications on the subject \cite{J30,W50,J29,J28}.

@@

  \llift{\exists} X_\mu \lambdot \phi X]}$$ (\texttt{v} in the THF0

@@

uppercase $P$, for `positive', from Fig.~\ref{fig1}).  The

@@

for example, shows that this theorem follows from axioms A2 and A1

@@

$\begin{array}{llllllll}

@@

 & & \text{A1}, \text{A2}, \text{D1}, \text{A3}, \text{A4}, \text{D2'}, \text{D3}, \text{A5} & \text{KB} & \text{UNS} & \text{---/---} & \text{---/---} & \text{---/---} \\ 

>>>>

In this paper we focus on the core aspect of our work related to AI: proof automation with HOL-ATPs (cf.~aspects (ii)--(iv)
above). % In addition, we present the detailed technical
% background, i.e., the embedding of \HOML in \HOL.  None of these
% aspects have been adressed in detail yet in any of our other existing 
% (short and partly non-reviewed) papers on the subject \cite{J30,W50,J29,J28}.
The particular contributions of this paper are as follows: 
% In Sec.~\ref{sec1} we present Scott's script \cite{ScottNotes} of
% G\"odel's proof, and we summarize the results of our
% recent automation of this script with HOL-ATPs. 
In Sec.~\ref{sec2} we present an elegant embedding of 
\HOML \cite{Gallin75,homl} in \HOL \cite{andrewsSEP,B5}. 
This background theory extends and
adapts the work as presented in \cite{J23,B9}.
In Sec.~\ref{sec3}, we present details on the
encoding of this embedding and of G{\"o}del's argument in the
concrete THF syntax \cite{J22} for \HOL, and we report on the
experiments we have conducted with HOL-ATPs.  The main findings of
these experiments are summarized in Sec.~\ref{sec4}.
% : in addition to
% automating G{\"o}del's proof script we have automated also some further
% results (e.g.~modal collapse), and we have studied
% the argument under modified logical settings (varying vs.~constant domain semantics).
% (by switching from constant
% domain semantics to varying domain semantics)
Related and future work is addressed in Sec.~\ref{sec5}, and the paper
is concluded in Sec.~\ref{sec6}.
%  where we also discuss potential
% implications towards a \emph{computational theoretical
%   philosophy}.
None of the above aspects have been addressed (at least
not in depth) in any of our other existing (short and partly
non-reviewed) publications on the subject \cite{J30,J29,W50,J28}.

@@

  \llift{\exists} X_\mu \lambdot \phi X]}$$ \texttt{v} in the THF0

@@

uppercase $P$, for `positive', from Fig.~\ref{fig1}.  The

@@

for example, shows that theorem T1 follows from axioms A2 and A1

@@

$\begin{array}{llllllll} \\[-.5em]

@@

 & & \text{A1}, \text{A2}, \text{D1}, \text{A3}, \text{A4}, \text{D2'}, \text{D3}, \text{A5} & \text{KB} & \text{UNS} & \text{---/---} & \text{---/---} & \text{---/---} \\ \\[-.5em]
